---
title: "R training part 2"
format: html
execute: 
  echo: fenced
Params: 
  org_code = "RF4"
---

## NHSRtheme

An R package that can provide themes to ggplot for producing charts that follow the NHS Identity [guidelines](https://www.england.nhs.uk/nhsidentity/identity-guidelines/).

## Formatting in HTML

### Figure size

Quarto has [default height & width](https://quarto.org/docs/computations/execution-options.html#figure-options) set, but you can override these using `fig-width` , `fig-height` , or `out-width`, `out-height`, among [many options](https://quarto.org/docs/reference/cells/cells-knitr.html#figures). Unit identifies that can be used include px, cm, and %. This can be done either in the YAML at the top of the file so that it is applied to all the figures produced, or in a code chunk so that it is applied only to that chunk.Execution options set in a code chunk will take precedence over those set in the YAML.

``` yaml
---
title: "My Document"
format: 
  html:
    fig-width: 8
    fig-height: 6
  pdf:
    fig-width: 7
    fig-height: 5
---
```

or

```{{r}}
#| out-width: 20%


```

### Tabs

You can introduce tabsets using `::: {.panel-tabset}` at the start of the section and end with `:::`. A tab will be added for each heading contained within -

::: panel-tabset
## Tab 1

```{r}
1+1
```

## Tab 2

```{r}
2+2
```
:::

## Saving & file locations

### {usethis} package

```{{r}}
install.packages("usethis")
usethis::edit_r_environ()                 # Opens the .Renviron file
```

Then set your path to the SharePoint folder you've sync'd to your OneDrive, e.g. `MY_PATH = "C:/Users/LyndaHoward/NHS England/Estates Support Office - ERIC Work Programme"`

Save and close the file. Restart R for the changes to take effect.

The variable `MY_PATH` can be retrieved using the `Sys.getenv()` function. You can add as many paths as you need in this way. file.path(Sys.getenv("MY_PATH"),"test")

## Parameters in Quarto & rmarkdown

### rendering with params

```{{r}}
quarto::quarto_render(
  input = "quarto_file.qmd",                        # Input Quarto file   
  output_format = "html",                           # Output format   
  output_file = "Report.html",                      # Output file name   
  execute_params = list(                            # Named list of parameters     
    trust = "RXR",    
    year = 2023   
    ))
```

## HELP!!!

When you need to post on a forum asking for help you'll often be asked to include a reprex (reproducible example). The dput() function allows you to create this.

```{r}
#create a small dataset
help_needed <- NHSRdatasets::ae_attendances |>  mutate(year = year(period)) |>  filter(year == 2019, between(breaches, 10, 100))|> filter(org_code %in% c('RF4', 'R1H', 'RYC'))

#create the code that will reproduce this dataset
dput(help_needed)
```
